Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2025-01-09 manuscript/manuscript.Rmd
Revised introduction

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle        : "Flattering Advice"
author: 
  - name          : "Amanda Chen"
    affiliation   : "1"
    corresponding : yes    
    email         : "zchengj@connect.ust.hk"
    address       : "Department of Management, The Hong Kong University of Science and Technology, Hong Kong"
  - name          : "David Hagmann"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "The Hong Kong University of Science and Technology"
note : |
  `r format(Sys.time(), "%e %B, %Y")`
abstract: |
  Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (n = 3,900), we demonstrate that advisers prioritize avoiding disappointment, even at the expense of accuracy. In Study 1, advisers financially rewarded for accuracy still tailored recommendations to aspirational goals expressed by advisees, resulting in poorer advice. When incentivized to be liked, advisers provided even more flattering advice, and advisees rated these advisers as more likable, despite the advice being less honest and less accurate (Study 2). In Study 3, advisers incorporated advisees' subjective and uninformative performance expectations into recommendations, leading to systematically different advice for men and women. Because men expected to perform better than women, they received more aspirational advice, even when their gender was unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making and equity.

keywords          : "Gender difference, Advice, Interpresonal relationship"
wordcount         : "6,104"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa7"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
classoption       : "man"
header-includes:
   - \usepackage{setspace, appendix, placeins, booktabs, tikz, siunitx}
   - \usepackage{tabularx, epigraph, float, colortbl, tabu, pdflscape}
   - \usepackage{tabularray}
   - \usetikzlibrary{positioning}
   - \raggedbottom
   - \usepackage[maxfloats=256]{morefloats}
   - \maxdeadcycles=1000
bibliography      : bibliography.bib
output: 
  # papaja::apa6_docx
  papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
always_allow_html: true
---

@@

# Introduction
Advice plays a critical role in both professional and personal decision-making, 
often serving as a crucial tool for navigating complex challenges and improving 
outcomes [@Gino2007; @Harvey1997; @Soll2009]. While prior research has 
extensively examined the *recipients* of advice (e.g., how they evaluate or act 
on recommendations) far less attention has been devoted to understanding the 
motivations and behaviors of the *givers* of advice. In this paper, we shift the 
focus to advisers, proposing that their recommendations are shaped not only by 
accuracy considerations but also by interpersonal concerns. Specifically, we 
propose that advisers factor in advisees' subjective expectations, providing 
flattering advice that aligns with these expectations, even when doing so leads 
to objectively worse recommendations. Notably, this tendency persists in our 
experimental settings, where advisees' expectations are clearly uninformative 
and advisers are financially incentivized for accuracy.

Advice is often viewed as a mechanism to improve decision satisfaction or 
prediction accuracy, assuming both parties prioritize solving the problem at 
hand. However, prior research highlights that interpersonal dynamics can 
complicate these interactions. Advice seekers may avoid asking for help due to 
fear of rejection [@Benabou2022], and people generally have been shown to avoid
information that could lead to disappointment [@Gill2012; @Golman2017]. 
They are motivated to avoid unpleasant truths in an ethical dilemma [@Shalvi2019] 
and dislike those who deliver unwelcome feedback even when they had no role in
the outcome [@John2019]. These tendencies align with models of belief utility 
[@Loewenstein2018], which suggest that individuals experience 
hedonic costs when receiving bad news. A natural reference point for evaluating 
the valence of advice is the advisee's expectation in the absence of advice
[@Koszegi2006]. Consequently, advisers who are aware of expectations may 
strategically adapt their advice to avoid delivering bad news, even at the 
expense of accuracy.

This interpersonal dimension of advice-giving has significant implications for 
understanding systematic gender differences in advice quality. Men are more 
likely than women to receive aspirational advice and encouragement to pursue 
competitive or risky opportunities [@Kanze2018]. Conversely, women are often 
shielded from criticism [@Jampol2023], which, while intended to be kind, 
may ultimately limit their professional growth. We propose a novel mechanism to 
explain these patterns: advisers calibrate their recommendations to match 
advisees’ expectations. Because, in many tasks (including those in our experiments)
men are more confident and have higher expectations about their performance than
do women, they are more likely to receive flattering (but, in our setting, worse) 
advice.

We present evidence for this mechanism from three preregistered and incentivized 
experiments involving real interactions between advisers and advisees (N = 3,900). 
In Study 1, we experimentally manipulate advisee's expectations and show
that advisers account for those expectations, even when they offer no relevant
insight for the decision. In Study 2, we show that advisers put more weight on
interpersonal concerns when they are incentivized to be liked. Advisees reward
this flattery by rating advisers as more likeable and equally trustworthy compared
to those offering more honest (and accurate) feedback. In Study 3, we measure
expected performance on a math quiz and show that advisers take these expectations
into account even when they have information about actual performance. Because
men expect to score higher, they are more likely to be advised to compete against
a stronger group for a higher potential payment than are women, even when advisers
are unaware of the advisees' gender. Notably, this leads to worse advice for men.

@@

We report all manipulations, measures, and data exclusion in this and the following study. The preregistration reports, screenshots of all experimental materials, and the analysis code to replicate all statistical analyses and figures are available on the Open Science Framework
(<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>).

# Study 1
We first examine whether advisers take into account the expectations of advisees
when providing a recommendation. A group of participants ("Advisees") complete 
a quiz consisting of ten questions that draw on ego-relevant domains. They then
express a non-binding preference to compete against a group of high performers
or low performers on the same task, based on their past performance. We nudge
participants towards picking either the low or the high performer group by 
selecting one of the two by default. A second group of participants ("Advisers") 
observe Advisees' score on the quiz and, in one treatment, also which group
they had selected. We predicted that Advisees who were nudged toward the high
performer group would be more likely to receive advice to compete against this
high group when Advisers observed their non-binding choice.

## Methods
In a preliminary stage, we recruited `r scales::comma(dim(study1_ability50)[1])` 
participants from Prolific and gave them five minutes to complete a quiz 
consisting of ten items. The quiz included word puzzles,
identifying emotions from photos, and selecting the best responses for hypothetical
scenarios. These questions were adapted from surveys that measure problem-solving,
emotional intelligence, and communication skills---which we pre-tested as being
important in modern society and hence where participants might have a stake in
doing well. Participants received a bonus of five cents for each correctly
answered question. We then ranked them based on their score and label the 20
participants with the lowest (highest) scores as the Low (High) Performer Group.

Next, we recruited `r scales::comma(dim(study1_time1)[1])` participants who
completed the same 10-item quiz, also earning five cents for each correct answer.
We then asked them to express a non-binding preference for whether they would
like to compete against the High or Low Performer Group. We informed them that
they would be invited back at a later date when they could make a binding decision
and they could earn a bonus based on whether their score on this quiz was equal
to or higher than a randomly selected member from their comparison group.
If they picked the High Performer Group and had an equal or higher score than
a randomly selected member of that group, they would earn a bonus of 50 cents.
If they picked the Low Performer Group, they could earn a bonus of 20 cents.
If their score was lower in this comparison, they would not earn an additional
bonus. We randomized which of the two groups (High or Low Performers) was 
selected by default, and participants were free to select the other group.
Participants also made a guess as to their score, and the survey concluded with 
basic demographic questions.

We then recruited `r scales::comma(dim(study1_time2_giver)[1])`
participants for the role of advisers, and the focal part of our experiment. 
They were informed of the ability quiz that participants had completed, how the 
Low Performer Group and High Performer Groups were constructed, as well as the
choice and incentives for the other participants. We then asked participants
to give advice to ten participants on which group they should compete against.
Advisers were randomly assigned to one of two treatments. In the
"Performance" treatment, they observed the score of the other participant.
In the "Performance + Expectation" treatment, they additionally observed the
advisee's non-binding choice for which group to compete against.
However, since the outcome was based solely on the past quiz score and advisers,
but not advisees, know the score, this initial choice was not informative.
Advisers could earn the same bonus as one of the participants they had given
advice to and who returned to make a decision. The survey concluded with basic 
demographic questions.

Finally, we invited advisees back for the follow-up survey. Following our 
preregistration, we kept the survey open for 7
days. In total, `r scales::comma(dim(study1_time3)[1])` participants
returned. The short survey reminded them of the task they
completed in the previous survey, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally reminded them of how
many questions they guessed they had answered correctly. Importantly,
they were not informed of their true score or the score of the groups
they could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

@@

We begin by examining the performance of the Stage 1 participants.
Because our default treatment takes place after participants completed the
ability quiz, we do not expect a difference in performance across
the Low and High Default treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study1_time1[study1_time1$default == "Low",]$Score),2)`
and
`r round(mean(study1_time1[study1_time1$default == "High",]$Score),2)`
for the Low Default and High Default treatments, respectively,
`r papaja::apa_print(t.test(Score ~ default, data = study1_time1, var.equal = T))$statistic`).
The default manipulation affected their initial choices. In the "Low Default" treatment, 
only `r round(mean(study1_time1[study1_time1$default == "Low",]$Guess),2)*100`% 
of participants initially preferred competing against the High Performer Group, 
compared to `r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`%
of participants who saw the High Performer Group highlighted ($\chi^2$(1, n = 201) = 36.62, p < .001).

@@

We now examine whether advisers took the advisees' expectations into account. 
Recall that advisees who were defaulted to competing against the High Performer 
Group were indeed more likely to choose this group, which, from the advisers' 
perspective, could signal high confidence in their abilities. Our theory 
predicts that showing expectations should lead to more flattering advice—
specifically, a recommendation to compete against the High Performer Group—
for advisees with a high default choice compared to those with a low 
default choice. Column 1 of Table \@ref(tab:study1regs) reports a linear
probability model with advice to compete against the high group
as the outcome measure, incorporating the default choice advisees saw, whether expectations
were shown to the adviser, and the interaction of the two. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
As predicted, we find a marginally significant interaction effect: advisees 
with a default choice to compete against the High Performer Group (nudged 
to have high expectations, though not necessarily holding them) were more 
likely to receive advice to compete against stronger competitors when 
their expectations were shown compared to when they were hidden.
We reported a linear probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee in Column 2 of Table \@ref(tab:study1regs).
We observe a significant main effect of the default treatment, where a high default
increased the likelihood of receiving flattering advice. It is worth noting that, 
since only `r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their default 
choice, this represents a conservative test of the effect of high expectations on 
the likelihood of receiving flattering advice.

@@

Next, we turn our attention to the advisers (Figure
\@ref(fig:study2advice)). In the Accuracy condition, those uploading
selfies were advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offered flattering advice.[^3] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers inferred that advising someone that they
were more attractive would make the adviser appear more likeable and
therefore provided advice that communicated a more favorable impression
of the participants' attractiveness. The distribution shown in Figure
\@ref(fig:study2advice) shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^3]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they had given and make a recommendation that combines
    their own belief with a uniform prior.

@@

Like before, advisers were randomly assigned to one of two treatments. In the
"Baseline" treatment, they only observed the score of the advisee on the
mathematics quiz. In the "Expectation" treatment, they observed the
score as well as the advisee's guess for how many questions they
answered correctly. We want to emphasize that sice the outcome is determined only by
the past score on the quiz, the advisee's guess is immaterial to which
group they should compete against. Moreover, in neither treatment did
they receive any demographic information about their advisee.
Participants gave recommendations to 10 advisees, which unbeknownst to
them were five men and five women matched to have identical performance
on the test.[^1] They were informed that if their advice was shown to a
participant who returned for the follow-up survey, they would receive
the identical bonus as that participant. The survey then concluded with
basic demographic questions.

[^1]: We made this decision to account for the possibility of gender
    differences in performance.

@@

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of Table
\@ref(tab:study3regs). Displaying expectations led men to be advised to
compete against the High Performer group more often, and this advise
turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^2]

[^2]: This analysis was not preregistered, and we note here that the
    reduction in expected earnings is small. However, it is interesting
    that expectations have a negative effect for men who underestimate
    their performance on average. One possibility is that advisers
    suggest the High Performer group more often than is optimal. We
    return to the possibility that advice is overall biased to be
    flattering in Study 2.

@@

```{=tex}
\begingroup
\singlespacing
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\noindent
```
::: {#refs}
:::

```{=tex}
\endgroup
```

>>>>

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle        : "Flattering Advice"
author: 
  - name          : "Amanda Chen"
    affiliation   : "1"
    corresponding : yes    
    email         : "zchengj@connect.ust.hk"
    address       : "Department of Management, The Hong Kong University of Science and Technology, Hong Kong"
  - name          : "David Hagmann"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "The Hong Kong University of Science and Technology"
note : |
  `r format(Sys.time(), "%e %B, %Y")`
abstract: |
  Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In Study 1, advisers financially rewarded for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in worse advice. When incentivized to be liked, advisers provide even more flattering advice, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate (Study 2). The desire to avoid disappointment may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

keywords          : "Gender difference, Advice, Interpresonal relationship"
wordcount         : "6,104"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa7"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
classoption       : "man"
header-includes:
   - \usepackage{setspace, appendix, placeins, booktabs, tikz, siunitx}
   - \usepackage{tabularx, epigraph, float, colortbl, tabu, pdflscape}
   - \usepackage{tabularray}
   - \usetikzlibrary{positioning}
   - \raggedbottom
   - \usepackage[maxfloats=256]{morefloats}
   - \maxdeadcycles=1000
bibliography      : bibliography.bib
output: 
  # papaja::apa6_docx
  papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
always_allow_html: true
# bibliography: references.bib
---

@@

# Introduction

Career trajectories for men and women often differ substantially, even
among those with similar qualifications. One source of such differences
may be the advice that people receive. Prior work has documented that
men receive more aspirational advice and women more risk-averse advice,
which may explain why women are less likely to apply for and ultimately
obtain more rewarding positions (Kanze et al., 2018). Research on gender
discrimination has focused on the role of unconscious bias (Greenwald et
al., 2009; Greenwald & Banaji, 1995), but interventions that try to
reduce this bias have been largely unsuccessful (Chang et al., 2019;
Paluck & Green, 2009). Recent work has proposed cognitive, rather than
unconscious, biases that can lead to gender discrimination through the
formation of false beliefs (Hagmann et al., 2024). Here, we propose a
novel source of differences in advice by drawing on recent work on
belief-based utility, expectation-based reference points (Köszegi, 2006;
Kőszegi & Rabin, 2009), and the hedonic consequences of information
(Golman et al., 2017).

Advice figures prominently in workplace and personal decision-making,
shaping choices about promotions, career paths, and everyday dilemmas
[@Gino2007; @Harvey1997; @Soll2009]. Its fundamental premise is to
improve outcomes by enabling decision-makers to learn from others'
expertise. However, while honest advice can lead to better decisions, it
can also lead to disappointment when it suggests a desired outcome is
unlikely to materialize. Advisers may be motivated to offer flattering
recommendations that align with advisees' hopes, sparing them the
disappointment of a more sobering forecast.

Advice with the potential for disappointment appears in diverse
organizational settings: supervisors deliver feedback on employee
performance, lawyers convey likely outcomes of litigation to their
clients, and academics offer thoughts on colleagues' manuscripts.
Ideally, honest information would help people identify their strengths
and weaknesses, make informed decisions about whether to pursue legal
cases, and strengthen papers prior to submission to academic journals,
thus improving long-term outcomes. In practice, however, interpersonal
considerations may prevent candor. When someone's expectations are (too)
high, an adviser might hesitate to deliver information that undermines
it---particularly if doing so risks negative interpersonal
repercussions.

People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves [@Golman2017; @Ho2018; @Loewenstein2018]. Moreover, they
often prefer advisers who withhold bad news from them [@Shalvi2019] and
punish those who communicate bad news [@John2019]. Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

In our experiments, however, advisees have no opportunity to punish the
adviser, and advisers' financial incentives are linked to the decisions
the advisees make. We propose that advisers nonetheless have reason to
provide flattering advise because they recognize the psychological toll
of disappointing news. As a result, they may similarly incur a hedonic
cost for delivering unpleasant information that they anticipate may
distress the advisee. Thus, even in the anonymized context of an online
experiment, we hypothesize (and find) that advisers are relucant to
disappoint advisees and advice is thus biased upward, away from accuracy
and toward confirming advisees' priors. We show that this reduces the
quality of advice—but that advisees nonetheless view flattering advisers
more favorably.

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their feedback to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.\^[In some cases, however, repeated
interaction might also increase incentives for honest feedback. Someone
who is known to persistently give overoptimistic advice may be viewed as
less trustworthy in the long run.]

In Study 1, we nudge advisees into a preference for competing against
either a group of top-performers or a group of low-performers, through
the use of a default option. Advisers who observe the actual performance
of advisees and therefore have the relevant information needed to make a
recommendation nonetheless take into account the initial preference when
recommending which group they should compete against. In Study 2, we ask
advisees to upload photos of themselves and advisers rank them on their
attractiveness. Advisers then recommend which rank the advisee should
bet on, receiving an incentive either for the advisee's accuracy or for
being evaluated favorably. Both groups of advisers recommend a rank that
is more attractive than what they themselves have evaluated the advisee,
and those incentivized for likability further inflate the ranking.
Advisees evaluate advisers more favorably when they recommend betting on
a more attractive rank, including viewing them as more trustworthy.

In Study 3, we examine a setting in which participants report their
expected performance on a mathematics quiz. Advisers observe the
test-taker's true score and, in one treatment, the score they guessed
they would receive. They then recommend whether the advisee should
compete against a group of high performers or low performers.
Importantly, this "competition" is based on the past performance
observed by the adviser and thus the advisee's expectation does not
provide instrumental information. However, we find that (1) men expect
to perform better than women, given identical performance, and (2)
advisers take these expectations into account. As a result, when
expectations (but not gender) are known to advisers, the advice given to
men and women differs, and men are more often advised to compete against
the group of high performers. Notably, this leads to worse advice for
men than for women.

@@

We report all manipulations, measures, and data exclusion in our
experiments. The preregistration reports, screenshots of all
experimental materials, and the analysis code to replicate all
statistical analyses and figures are available on the Open Science
Framework
(<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>).

# Study 1

We first examine whether advisers take into account the expectations of
advisees when providing a recommendation. A group of participants
("Advisees") complete a quiz consisting of ten questions that draw on
ego-relevant domains. They then express a non-binding preference to
compete against a group of high performers or low performers on the same
task, based on their past performance. We nudge participants towards
picking either the low or the high performer group by selecting one of
the two by default. A second group of participants ("Advisers") observe
Advisees' score on the quiz and, in one treatment, also which group they
had selected. We predicted that Advisees who were nudged toward the high
performer group would be more likely to receive advice to compete
against this high group when Advisers observed their non-binding choice.

In a preliminary stage, we recruited
`r scales::comma(dim(study1_ability50)[1])` participants from Prolific
and gave them five minutes to complete a quiz consisting of ten items.
The quiz included word puzzles, identifying emotions from photos, and
selecting the best responses for hypothetical scenarios. These questions
were adapted from surveys that measure problem-solving, emotional
intelligence, and communication skills---which we pre-tested as being
important in modern society and hence where participants might have a
stake in doing well. Participants received a bonus of five cents for
each correctly answered question. We then ranked them based on their
score and label the 20 participants with the lowest (highest) scores as
the Low (High) Performer Group.

Next, we recruited `r scales::comma(dim(study1_time1)[1])` participants
for the role of Advisees. They completed the same 10-item quiz, also
earning five cents for each correct answer. We then asked them to
express a non-binding preference for whether they would like to compete
against the High or Low Performer Group. We informed them that they
would be invited back at a later date when they could make a binding
decision and they could earn a bonus based on whether their score on
this quiz was equal to or higher than a randomly selected member from
their comparison group. If they picked the High Performer Group and had
an equal or higher score than a randomly selected member of that group,
they would earn a bonus of 50 cents. If they picked the Low Performer
Group and outperformed a randomly selected member of that group, they
could earn a bonus of 20 cents. If their score was lower in this
comparison, they would not earn an additional bonus. We randomized which
of the two groups (High or Low Performers) was selected by default, and
Advisees were free to select the other group. They also made a guess as
to their score, and the survey concluded with basic demographic
questions.

We then recruited `r scales::comma(dim(study1_time2_giver)[1])`
participants for the role of Advisers, and the focal part of our
experiment. They were informed of the ability quiz that the Advisees had
completed, how the Low Performer Group and High Performer Groups were
constructed, as well as the choice and incentives for the Advisees. We
then asked Advisers to give advice to ten participants on which group
they should compete against. Advisers were randomly assigned to one of
two treatments. In the "Performance" treatment, they observed only the
score of the other participant. In the "Performance + Expectation"
treatment, they additionally observed the advisee's non-binding choice
of which group to compete against. However, since the outcome was based
solely on the past quiz score and advisers, but not advisees, know the
score, this initial choice was not informative for the purpose of
recommending competing against one group or the other. Advisers could
earn the same bonus as one of the participants they had given advice to
and who returned to make a decision. The survey concluded with basic
demographic questions.

Finally, we invited Advisees back for the follow-up survey. Following
our preregistration, we kept the survey open for 7 days. In total,
`r scales::comma(dim(study1_time3)[1])` Advisees returned. The short
survey reminded them of the task they completed in the previous survey,
informed them that other participants from Prolific had observed their
real score and given them advice on which group to compete against, and
finally reminded them of how many questions they guessed they had
answered correctly. They were not informed of their true score or the
score of the groups they could compete against. Participants then
observed the advice from a randomly selected adviser and made their
decision.

@@

We begin by examining the performance of the Stage 1 participants.
Because our default treatment takes place after participants completed
the ability quiz, we do not expect a difference in performance across
the Low and High Default treatments. Indeed, the two groups scored no
different from one another
(`r round(mean(study1_time1[study1_time1$default == "Low",]$Score),2)`
and
`r round(mean(study1_time1[study1_time1$default == "High",]$Score),2)`
for the Low Default and High Default treatments, respectively,
`r papaja::apa_print(t.test(Score ~ default, data = study1_time1, var.equal = T))$statistic`).
The default manipulation affected their initial choices. In the "Low
Default" treatment, only
`r round(mean(study1_time1[study1_time1$default == "Low",]$Guess),2)*100`%
of participants initially preferred competing against the High Performer
Group, compared to
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`%
of participants who saw the High Performer Group highlighted
($\chi^2$(1, n = 201) = 36.62, p \< .001).

@@

We now examine whether advisers took the advisees' expectations into
account. Recall that advisees who were defaulted to competing against
the High Performer Group were indeed more likely to choose this group,
which, from the advisers' perspective, could signal high confidence in
their abilities. Our theory predicts that showing expectations should
lead to more flattering advice--- specifically, a recommendation to
compete against the High Performer Group--- for advisees with a high
default choice compared to those with a low default choice. Column 1 of
Table \@ref(tab:study1regs) reports a linear probability model with
advice to compete against the high group as the outcome measure,
incorporating the default choice advisees saw, whether expectations were
shown to the adviser, and the interaction of the two. Because each
adviser made ten recommendations, we cluster standard errors at the
adviser level. As predicted, we find a marginally significant
interaction effect: advisees with a default choice to compete against
the High Performer Group (nudged to have high expectations, though not
necessarily holding them) were more likely to receive advice to compete
against stronger competitors when their expectations were shown compared
to when they were hidden. We reported a linear probability model on
advising to compete against the High Performer group for advisers in the
"Expectations" treatment, controlling for the true performance of the
advisee in Column 2 of Table \@ref(tab:study1regs). We observe a
significant main effect of the default treatment, where a high default
increased the likelihood of receiving flattering advice. It is worth
noting that, since only
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their
default choice, this represents a conservative test of the effect of
high expectations on the likelihood of receiving flattering advice.

@@

Next, we turn our attention to the advisers (Figure
\@ref(fig:study2advice)). In the Accuracy condition, those uploading
selfies were advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offered flattering advice.[^1] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers inferred that advising someone that they
were more attractive would make the adviser appear more likeable and
therefore provided advice that communicated a more favorable impression
of the participants' attractiveness. The distribution shown in Figure
\@ref(fig:study2advice) shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^1]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they had given and make a recommendation that combines
    their own belief with a uniform prior.

@@

Like before, advisers were randomly assigned to one of two treatments.
In the "Baseline" treatment, they only observed the score of the advisee
on the mathematics quiz. In the "Expectation" treatment, they observed
the score as well as the advisee's guess for how many questions they
answered correctly. We want to emphasize that sice the outcome is
determined only by the past score on the quiz, the advisee's guess is
immaterial to which group they should compete against. Moreover, in
neither treatment did they receive any demographic information about
their advisee. Participants gave recommendations to 10 advisees, which
unbeknownst to them were five men and five women matched to have
identical performance on the test.[^2] They were informed that if their
advice was shown to a participant who returned for the follow-up survey,
they would receive the identical bonus as that participant. The survey
then concluded with basic demographic questions.

[^2]: We made this decision to account for the possibility of gender
    differences in performance.

@@

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of Table
\@ref(tab:study3regs). Displaying expectations led men to be advised to
compete against the High Performer group more often, and this advise
turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^3]

[^3]: This analysis was not preregistered, and we note here that the
    reduction in expected earnings is small. However, it is interesting
    that expectations have a negative effect for men who underestimate
    their performance on average. One possibility is that advisers
    suggest the High Performer group more often than is optimal. We
    return to the possibility that advice is overall biased to be
    flattering in Study 2.

@@

\begingroup
\singlespacing
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}

\noindent

::: {#refs}
:::

\endgroup
